%% ════════════════════════════════════════════════════════════════════════════════
%% Academic Poster — Tree of Thoughts for Automated Code Debugging
%% CS 4782, Cornell University
%% Size: 36" × 24" landscape
%%
%% Compile: pdflatex poster.tex
%% Requires: beamerposter package  (tlmgr install beamerposter)
%% ════════════════════════════════════════════════════════════════════════════════
\documentclass[final]{beamer}

\usepackage[
  orientation=landscape,
  size=custom,
  width=91.44,   % 36 inches in cm
  height=60.96,  % 24 inches in cm
  scale=1.4
]{beamerposter}

\usepackage{lmodern}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning,fit}
\usepackage{tcolorbox}
\tcbuselibrary{skins,breakable}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{multicol}

%% ── Color palette ─────────────────────────────────────────────────────────────
\definecolor{cornellred}{RGB}{179,27,27}
\definecolor{lightgray}{RGB}{245,245,245}
\definecolor{darkgray}{RGB}{80,80,80}
\definecolor{accentblue}{RGB}{30,90,180}
\definecolor{accentgreen}{RGB}{40,150,80}
\definecolor{accentorange}{RGB}{220,100,20}

%% ── Beamer theme ──────────────────────────────────────────────────────────────
\usetheme{default}
\setbeamercolor{background canvas}{bg=white}
\setbeamercolor{headline}{fg=white,bg=cornellred}
\setbeamercolor{block title}{fg=white,bg=accentblue}
\setbeamercolor{block body}{fg=black,bg=lightgray}
\setbeamerfont{block title}{size=\large,series=\bfseries}
\setbeamerfont{block body}{size=\normalsize}

%% ── tcolorbox styles ──────────────────────────────────────────────────────────
\tcbset{
  mybox/.style={
    enhanced, breakable,
    colback=lightgray, colframe=accentblue,
    fonttitle=\bfseries\large,
    arc=4pt, boxrule=1.5pt,
    top=6pt, bottom=6pt, left=8pt, right=8pt,
    title=#1
  },
  redbox/.style={
    enhanced,
    colback=cornellred!8, colframe=cornellred,
    fonttitle=\bfseries\large,
    arc=4pt, boxrule=1.5pt,
    top=6pt, bottom=6pt, left=8pt, right=8pt,
    title=#1
  },
  greenbox/.style={
    enhanced,
    colback=accentgreen!8, colframe=accentgreen,
    fonttitle=\bfseries\large,
    arc=4pt, boxrule=1.5pt,
    top=6pt, bottom=6pt, left=8pt, right=8pt,
    title=#1
  },
}

\setlist[itemize]{leftmargin=1.4em, itemsep=3pt, topsep=3pt}

%% ════════════════════════════════════════════════════════════════════════════════
\begin{document}
\begin{frame}[t]{}

%% ── HEADER ────────────────────────────────────────────────────────────────────
{\usebeamercolor[bg]{headline}\color{white}
\begin{tikzpicture}[remember picture, overlay]
  \fill[cornellred] (current page.north west) rectangle
    ([yshift=-3.8cm]current page.north east);
\end{tikzpicture}
\vspace{0.3cm}
\centering
{\fontsize{38}{44}\bfseries\selectfont
  Tree of Thoughts for Automated Code Debugging\par}
\vspace{0.3cm}
{\fontsize{22}{28}\selectfont
  Ronald Feng \quad Cornell University \quad CS 4782: Intro to Deep Learning\par}
\vspace{0.5cm}
}

%% ── THREE-COLUMN BODY ────────────────────────────────────────────────────────
\vspace{0.6cm}
\begin{columns}[t,totalwidth=\textwidth]

%% ══════════════════════ COLUMN 1 ══════════════════════════════════════════════
\begin{column}{0.32\textwidth}

\begin{tcolorbox}[mybox={Introduction \& Motivation}]
LLMs excel at coding tasks but struggle with \textbf{debugging}, which requires:
\begin{itemize}
  \item Exploring \emph{multiple} hypotheses about the root cause
  \item Evaluating each hypothesis against test evidence
  \item \textbf{Backtracking} when a diagnosis is wrong
\end{itemize}

\medskip
\textbf{Standard IO / CoT prompting} commits to a single reasoning path—it cannot recover from an incorrect initial diagnosis.

\medskip
\textbf{Tree of Thoughts (ToT)} \textit{(Yao et al., NeurIPS 2023)} generalizes CoT by allowing exploration of a \textbf{tree of intermediate reasoning steps} using BFS or DFS search.

\medskip
\textcolor{accentblue}{\textbf{Our contribution:}} Apply ToT to Python bug fixing. Code debugging is uniquely suited because \textbf{test execution provides an unambiguous evaluation signal}—no LLM self-judgment needed.
\end{tcolorbox}

\vspace{0.4cm}

\begin{tcolorbox}[mybox={Original Paper: Key Result}]
\centering
\begin{tabular}{lc}
\toprule
\textbf{Method} & \textbf{Game of 24 Success} \\
\midrule
IO              & 4\%  \\
CoT             & 4\%  \\
\textbf{ToT-BFS} & \textbf{74\%} \\
\bottomrule
\end{tabular}

\medskip
\textit{Table 1 of Yao et al. (2023): ToT yields an 18× improvement via structured search.}

\medskip
\textbf{Hypothesis:} The same structured-search advantage appears in code debugging, where the search space is the set of possible root-cause diagnoses.
\end{tcolorbox}

\vspace{0.4cm}

\begin{tcolorbox}[mybox={Dataset}]
\textbf{HumanEval-Bugs} (50 Python problems)
\begin{itemize}
  \item Derived from OpenAI HumanEval
  \item Systematic mutations per category:
\end{itemize}

\medskip
\begin{tabular}{lc}
\toprule
\textbf{Bug Type} & \textbf{Count} \\
\midrule
\texttt{off\_by\_one}       & 15 \\
\texttt{wrong\_operator}    & 15 \\
\texttt{incorrect\_return}  & 11 \\
\texttt{missing\_condition} & 9 \\
\midrule
Total                        & 50 \\
\bottomrule
\end{tabular}

\medskip
Ground-truth: original HumanEval test assertions executed in sandboxed subprocess (5 s timeout).
\end{tcolorbox}

\end{column}

%% ══════════════════════ COLUMN 2 ══════════════════════════════════════════════
\begin{column}{0.35\textwidth}

\begin{tcolorbox}[mybox={Methodology: ToT Debugging Pipeline}]

% Tree diagram
\centering
\begin{tikzpicture}[
  node distance=0.8cm and 1.0cm,
  every node/.style={font=\small},
  state/.style={rectangle, rounded corners, draw=accentblue, fill=accentblue!12,
                text width=3.5cm, align=center, minimum height=0.7cm},
  hyp/.style={rectangle, rounded corners, draw=accentorange, fill=accentorange!12,
              text width=2.8cm, align=center, minimum height=0.7cm},
  fix/.style={rectangle, rounded corners, draw=accentgreen, fill=accentgreen!12,
              text width=2.6cm, align=center, minimum height=0.7cm},
  arrow/.style={->, >=stealth, thick}
]
  \node[state] (root) {Buggy Code\\+ Test Failures};

  \node[hyp, below left=0.8cm and 1.6cm of root] (h1) {Hypothesis 1\\(scored)};
  \node[hyp, below=0.8cm of root]                (h2) {Hypothesis 2\\(scored)};
  \node[hyp, below right=0.8cm and 1.6cm of root](h3) {Hypothesis 3\\(scored)};

  \node[fix, below left=0.8cm and 0.6cm of h1]   (f1) {Fix 1a};
  \node[fix, below right=0.8cm and 0.0cm of h1]  (f2) {Fix 1b};
  \node[fix, below=0.8cm of h2]                  (f3) {Fix 2a};
  \node[fix, below left=0.8cm and 0.0cm of h3]   (f4) {Fix 3a};
  \node[fix, below right=0.8cm and 0.6cm of h3]  (f5) {Fix 3b};

  \draw[arrow] (root) -- (h1);
  \draw[arrow] (root) -- (h2);
  \draw[arrow] (root) -- (h3);
  \draw[arrow] (h1) -- (f1);
  \draw[arrow] (h1) -- (f2);
  \draw[arrow] (h2) -- (f3);
  \draw[arrow] (h3) -- (f4);
  \draw[arrow] (h3) -- (f5);

  % Mark one fix as passing
  \node[fix, draw=accentgreen!80!black, fill=accentgreen!30, below right=0.8cm and 0.0cm of h3, minimum height=0.7cm] (f5pass) {Fix 3b \checkmark};
\end{tikzpicture}

\medskip\raggedright
\textbf{Four components:}
\begin{enumerate}[noitemsep]
  \item \textbf{Thought Generator} — GPT-4o-mini produces $k{=}3$ numbered bug hypotheses from buggy code + failing tests
  \item \textbf{Thought Evaluator} — Hybrid: 60\% LLM score (1–10) + 40\% execution signal (quick-fix + test run)
  \item \textbf{Search Controller} — BFS: expand all top-$k$ hypotheses; DFS: greedy + backtrack
  \item \textbf{Fix Generator} — For each selected diagnosis, generate $k$ complete corrected functions; verify via test execution
\end{enumerate}

\textbf{Base LLM:} GPT-4o-mini \quad \textbf{Cost:} $\sim$\$1–2 for 50 problems
\end{tcolorbox}

\vspace{0.4cm}

\begin{tcolorbox}[mybox={BFS vs. DFS Search}]
\begin{tabular}{p{4cm}p{5cm}p{5cm}}
\toprule
 & \textbf{BFS} & \textbf{DFS} \\
\midrule
\textbf{Strategy} & Expand all $k$ hypotheses before fixing & Commit to best hypothesis; backtrack on failure \\
\textbf{Fix Rate}  & 76\% & 74\% \\
\textbf{Avg Tokens}& 4,102 & 3,248 \\
\textbf{Backtracks}& 0 & 1.42 avg \\
\textbf{Best for}  & Max accuracy & Token efficiency \\
\bottomrule
\end{tabular}
\end{tcolorbox}

\end{column}

%% ══════════════════════ COLUMN 3 ══════════════════════════════════════════════
\begin{column}{0.30\textwidth}

\begin{tcolorbox}[greenbox={Results}]
\centering
\begin{tabular}{lcc}
\toprule
\textbf{Method} & \textbf{Fix Rate} & \textbf{Avg Tokens} \\
\midrule
IO              & 52\%  & 312 \\
CoT             & 60\%  & 590 \\
CoT-SC (n=5)    & 66\%  & 2,949 \\
ToT-DFS ($k=3$) & 74\%  & 3,248 \\
\textbf{ToT-BFS ($k=3$)} & \textbf{76\%} & \textbf{4,102} \\
\bottomrule
\end{tabular}

\medskip
\raggedright
\textbf{+24 pp} over IO \quad \textbf{+16 pp} over CoT

Mirrors the original ToT paper's large improvement over baselines.
\end{tcolorbox}

\vspace{0.4cm}

\begin{tcolorbox}[mybox={Fix Rate by Bug Type (ToT-BFS)}]
\centering
\begin{tabular}{lcc}
\toprule
\textbf{Bug Type} & \textbf{Fix Rate} & \textbf{Why?} \\
\midrule
\texttt{off\_by\_one}       & 87\% & Localized; hypothesis well-constrained \\
\texttt{wrong\_operator}    & 80\% & Single token change; easy to enumerate \\
\texttt{incorrect\_return}  & 73\% & Identifiable from output mismatch \\
\texttt{missing\_condition} & 56\% & Requires reasoning about absent code \\
\bottomrule
\end{tabular}
\end{tcolorbox}

\vspace{0.4cm}

\begin{tcolorbox}[redbox={Key Insights}]
\begin{itemize}
  \item \textbf{Execution beats self-eval}: Test pass/fail is a far stronger signal than LLM confidence ratings for evaluating fixes.

  \item \textbf{Backtracking matters}: DFS shows avg 1.42 backtracks/problem—confirming that single-path reasoning regularly fails.

  \item \textbf{Hypothesis quality is the bottleneck}: Missing-condition bugs (56\%) fail because the diagnostic hypothesis space is harder to enumerate—the LLM must reason about code that \emph{isn't} there.

  \item \textbf{Cost tradeoff}: ToT uses 7–13× more tokens than CoT. CoT-SC (66\%) is the best mid-cost option.
\end{itemize}
\end{tcolorbox}

\vspace{0.4cm}

\begin{tcolorbox}[mybox={Conclusion \& Future Work}]
\textbf{Conclusion:} ToT-BFS achieves \textbf{76\% fix rate} on HumanEval-Bugs vs.\ 52\% for IO and 60\% for CoT, demonstrating that structured tree search transfers successfully from puzzle tasks to automated code debugging.

\medskip
\textbf{Future directions:}
\begin{itemize}
  \item Dynamic analysis: pass stack traces and variable values to the hypothesis generator
  \item Larger $k$ or iterative refinement (re-rank with execution before fix step)
  \item Multi-file / repository-level debugging with retrieval-augmented hypothesis generation
  \item Evaluation on DebugBench across Python, Java, C++
\end{itemize}

\medskip
\textbf{References:}\\
\small
[1] Yao et al., \textit{Tree of Thoughts}, NeurIPS 2023.\\
[2] Chen et al., \textit{Evaluating LLMs on Code}, arXiv 2021.\\
[3] Wang et al., \textit{Self-Consistency}, ICLR 2023.
\end{tcolorbox}

\end{column}
\end{columns}

\end{frame}
\end{document}
